---
title: "King County Covid Cases"
author: "Beau Raines"
date: "7/4/2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(zoo)
library(hrbrthemes)
library(plotly)
library(emojifont)


```


```{r daily-and-rolling-14-day-average-cases, include = FALSE}
fips = read.csv("https://www2.census.gov/geo/docs/reference/codes/files/national_county.txt", header=FALSE)
nytimes= read.csv("https://github.com/nytimes/covid-19-data/blob/master/us-counties.csv?raw=true")
king_county = subset(nytimes,state =='Washington' & county == "King")
county_population = read.csv("https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv")

data <- data.frame(date = as.Date(tail(king_county$date,-1)), new_cases = diff(king_county$cases), new_deaths = diff(king_county$deaths))
## missing very very first day
p <- ggplot(data, aes(x=date, y=new_cases)) +
  # geom_line() +
  geom_bar(stat = "identity") +
  xlab("")+ylab("New Cases")
p+scale_x_date(date_labels = "%b-%y",date_breaks = "2 month")
p

x<-zoo(data$new_cases)

rolling_average = data.frame(date = as.Date(tail(king_county$date,-14)), rolling_new_cases_14d = rollmean(x,14))

p2 = ggplot(data=rolling_average, aes(x=date,y=rolling_new_cases_14d,group=1))+geom_line()+xlab("")
p2 +ylab("14 Day Rolling Average New Cases")
p2 +scale_x_date(date_labels = "%b-%y",date_breaks = "2 month")

combined_plot = ggplot() +
  geom_bar(stat="identity",data=data, aes(x=date, y=new_cases),fill="#FF9999", colour="#FF9999") +
  geom_line(data=rolling_average, colour= "darkblue", aes(x=date,y=rolling_new_cases_14d,group=1)) +
  scale_x_date(date_labels = "%b-%y",date_breaks = "2 month") +
  xlab("") + ylab("Cases")
combined_plot

# Add annotation
#combined_plot +
#  annotate(geom="text",x =as.Date('2020-12-15'),subset(data,date=='2020-12-15')$new_cases,label="Christmas Day") +
#  annotate(geom="point",x =as.Date('2020-12-15'),subset(data,date=='2020-12-15')$new_cases, size=10, shape=21, fill="transparent")

## Add titles
combined_plot = combined_plot +
  labs(title="King County Covid-19 Cases",caption="Data sourced from NY Times GitHub")

cases_per_100K_14d = sum(tail(data$new_cases,14))/subset(county_population, STNAME == 'Washington' & CTYNAME == "King County" ,select=POPESTIMATE2019)*100000

if(cases_per_100K_14d >= 75) {
  risk = "High"
} else if (cases_per_100K_14d >= 25 ){
  risk = "Moderate"
} else {
  risk = "Low"
}


## Write on plot
# combined_plot = combined_plot + annotate(geom="text",x=as.Date(Sys.Date()-21),y=2000,label=paste("Cases per 100k pop:",round(cases_per_100K_14d,2),sep="\n"))

# ipsum Theme
combined_plot = combined_plot  + 
  theme_ipsum()
combined_plot

## Source runs, but doesn't display the plots.
# it seems you have to create them as an object then display it afterwards
source("hospitalizations.R", local = knitr::knit_global())
# beds
```

## Notice

This report continues to be updated, but is no longer under active development. Please refer to the [Covid Dashboard](CovidDashboard.html) as it is the recipient of active development.


## King County Covid Cases

Yesterday, `r tail(data$date,1)` there were `r tail(data$new_cases,1)` new cases reported. King County has a population of `r format(subset(county_population, STNAME == 'Washington' & CTYNAME == "King County" ,select=POPESTIMATE2019),big.mark = ",")`. This is `r format(cases_per_100K_14d,digits=3,decimal.mark=".",  nsmall =1)` cases per 100k population over the last 14 days, which is `r risk` risk.

### new Cases
```{r valuebox, echo=FALSE, message = FALSE, warning=FALSE,fig.height=2, fig.align='left'}
yesterday_new_cases <- tail(data$new_cases,1)

df <- data.frame(
  x = 1,
  y = 1,
  h = 4.25,
  w = 6.25,
  value = c(yesterday_new_cases),
  info = c("new cases"),
  icon = fontawesome(search_fontawesome("chart"))[1],
  font_family = c("fontawesome-webfont",
                  "EmojiOne"),
  color = factor(1)
)

ggplot(df, aes(x, y, height = h, width = w, label = info)) +
  ## Create the tiles using the `color` column
  geom_tile(aes(fill = color)) +
  ## Add the numeric values as text in `value` column
  geom_text(color = "white", fontface = "bold", size = 10,
            aes(label = value, x = x - 2.9, y = y + 1), hjust = 0) +
  ## Add the labels for each box stored in the `info` column
  geom_text(color = "white", fontface = "bold",
            aes(label = info, x = x - 2.9, y = y - 1), hjust = 0) +
  coord_fixed() +
  scale_fill_brewer(type = "qual",palette = "Dark2") +
  ## Use `geom_text()` to add the icons by specifying the unicode symbol.
  geom_text(size = 20, aes(label = icon, family = font_family,
                           x = x + 1.5, y = y + 0.5), alpha = 0.25) +
  theme_void() +
  guides(fill = FALSE)



fig <- plot_ly(
  domain = list(x = c(0, 1), y = c(0, 1)),
  value = cases_per_100K_14d[,1],
  title = list(text = "Cases per 100k Population"),
  type = "indicator",
  mode = "gauge+number+delta",
  # delta = list(reference=52),
  gauge = list(
    bar = list(
      color = "gray"
    ),
    axis = list(
      range = c(0,150),
      dtick = 10
    ),
    steps = list(
      list(range = c(0,25),color = "lightgreen"),
      list(range = c(25,75), color = "lightyellow"),
      list(range = c(75,600), color = "red")
    )
  )) 
fig <- fig %>%
  layout(margin = list(l=20,r=30))

fig
```



In the last 7 days, there were `r sum(tail(data$new_cases,7))` new cases and `r last_7_days_reported_deaths` deaths.
```{r combined_plot, echo=FALSE}
ggplotly(combined_plot)  %>%
  layout(
    xaxis = list(
      rangeslider = list(type = "date"))
  )

```
```{r King County Last 14 Day Table, echo=FALSE, message = FALSE}
king_county_daily_cases <- tibble(Date = format(as.Date(tail(king_county$date,-1)),'%b %d'), `New Cases` = diff(king_county$cases), `New Deaths`=diff(king_county$deaths))
knitr::kable(king_county_daily_cases %>%
               tail(14) %>% t(),
             caption="King County 14 day History")


```

# Risk Factors

```{r King County Risk Factors, echo=FALSE, message = FALSE, warning=FALSE }
risk_data <- tibble(date = data$date,
              new_cases = data$new_cases,
              last_14_day_cases = NULL,
              last_14_day_cases_per_100k_pop = NULL
         )

for (i in 1:nrow(risk_data)) {
  startdate = risk_data[i,'date']$date
  enddate = risk_data[i,'date']$date -14
  risk_data$last_14_day_cases[i] <- risk_data %>%
                                subset(date <= startdate & date >= enddate) %>%
                                select(new_cases) %>% 
                                sum(na.rm=TRUE)
  risk_data$last_14_day_cases_per_100k_pop[i] <- risk_data$last_14_day_cases[i] / 2252782 * 100000
  if(risk_data$last_14_day_cases_per_100k_pop[i] >= 75) {
    risk_data$risk[i] = "High"
  } else if (cases_per_100K_14d >= 25 ){
    risk_data$risk[i] = "Moderate"
  } else {
    risk_data$risk[i] = "Low"
  }
}

ggplot(data = risk_data, mapping=aes(x=date,y=last_14_day_cases_per_100k_pop)) +
  geom_area(fill="lightgray",aes()) +
  geom_hline(yintercept=25,linetype="dashed",color="yellow")+
  geom_hline(yintercept=75,linetype="dashed",color="red") +
  labs(title="King County 14 Day Risk per 100k population",
       subtitle="Using Washington state risk factors: Less than 25, low, 25 - 75 medium, greater than 75, high. Effective June 30, 2021, Washington state lifted their mask mandate and switched to the CDC Risk factors.",
       caption="") +
  xlab("") +
  ylab("Cases per 100k population in last 14 days") +
  scale_x_date(date_labels = "%b-%y",date_breaks = "2 month")
  
```

## County Level Risk

```{r Washington county map, echo=FALSE, message = FALSE }
outlist = list()

washington = subset(nytimes,state =='Washington')

#Remove unknown county
washington <- subset(washington,county != 'Unknown')
nytimes <- subset(nytimes,county != 'Unknown')


for (c in unique(washington$county)) {
  county_info <-subset(washington,county == c & county != 'Unknown')
  county_fips = unique(county_info$fips)
  fooSTATE = as.numeric(substr(county_fips, 1, 2))
  fooCOUNTY = as.numeric(substr(county_fips, nchar(county_fips)-3+1, nchar(county_fips)))
  population = subset(county_population,STATE == fooSTATE & COUNTY == fooCOUNTY,select=c('POPESTIMATE2019'),na=0)$POPESTIMATE2019
  first_case_reported <- min(county_info$date)
  new_cases =data.frame(date = tail(county_info$date,-1), new_cases = diff(county_info$cases))
  last_14d_cases <- sum(subset(new_cases,date >= Sys.Date()-14)$new_cases)
  last_7d_cases <- sum(subset(new_cases,date >= Sys.Date()-7)$new_cases)
  out <- data.frame(state = unique(county_info$state),
                    county = unique(county_info$county),
                    county_fips = county_fips,
                    region = unique(county_info$fips),
                    first_case_reported = first_case_reported,
                    last_14d_cases = last_14d_cases,
                    last_7d_cases = last_7d_cases,
                    population = coalesce(population,0),
                    last_14d_cases_per_100k = coalesce(last_14d_cases / population *100000,0) ,
                    last_7d_cases_per_100k = coalesce(last_7d_cases / population *100000,0))
  outlist[[unique(county_info$fips)]] <- out
}



metrics = do.call(rbind,outlist)


library(choroplethr)
metrics$value = metrics$last_7d_cases_per_100k

choroplethr::county_choropleth(metrics,title="Last 7 Days Cases per 100k population",num_colors = 4,state_zoom = 'washington')+ scale_fill_brewer(palette=7)  

```

```{r Washington Last 14 Day Table, echo=FALSE, message = FALSE}
last_14_day_cases <- nytimes %>%
filter( date >= Sys.Date() - 15) %>%
group_by(state, county, fips) %>%
summarise(last_14_days_cases = sum(diff(cases)))
last_14_day_cases = left_join(last_14_day_cases,county_population,by=c("state"="STNAME","county"= "CTYNAME"))
last_14_day_cases = subset(last_14_day_cases, select = c("state","county","fips","last_14_days_cases","POPESTIMATE2019"))
last_14_day_cases$cases_per_100K_14d = last_14_day_cases$last_14_days_cases / last_14_day_cases$POPESTIMATE2019 * 100000

knitr::kable(subset(last_14_day_cases,state=='Washington',select=c('county','last_14_days_cases')),caption="Cases by County, Last 14 Days")
```

### County New Case History
```{r County New Case History, echo=FALSE, fig.align="center", fig.height=6, fig.width=10, message=FALSE, warning=FALSE, results=FALSE}
library(hrbrthemes)

#Remove unknown county
washington <- subset(washington,county != 'Unknown')
nytimes <- subset(nytimes,county != 'Unknown')

county_data = tibble()

for (c in unique(washington$county)) {
  county_info <-subset(washington,county == c & county != 'Unknown')
  county_fips = unique(county_info$fips)
  county_data <- bind_rows(county_data,tibble(state ='Washington',
                              county = c,
                              date = tail(county_info$date,-1),
                              new_cases = diff(county_info$cases),
                              new_deaths = diff(county_info$deaths) )
  )
}

county_data$date <- as.Date(county_data$date)

county_data %>% 
  # filter(date >='2021-01-01') %>%
  ggplot(aes(x=date,y=new_cases))+
  geom_line(group=1) +
  facet_wrap(~county,scales = "free") +
  labs(x="", y="",
       title="New Cases by county") + 
  theme_ipsum()
```
